\textbf{Supplementary Material}\\ \vspace{2cm} What is long-distance dispersal? And a taxonomy of dispersal events \\

### 2016-10-10 Manuscript/MS- What is LDD.tex
Edited refs and Suppl Mat.

The movement of pollen and seeds by animals and its consequences have intrigued population geneticists and field ecologists since the infancy of both research disciplines. Each has generated an impressive body of theoretical and empirical research through the past decades, yet advances have long been co-existing in ‘parallel worlds’ and the great synergistic potential of population genetics and demography for the study of plant dispersal by animals remains little explored. Knowledge gaps still having the imprint of this conceptual disconnection include the idea of long distance dispersal, and the paradoxes of forest fragmentation effects on genetic diversity \citep{Kramer:2008kg}, survival and persistence of relict tree species \citep{Hampe:2011bv}, rapid post-Pleistocene recolonization of vast continental areas in response to climate modification \citep{Clark:1998aa,Clark:1998vi}, among other persisting issues. This conceptual isolation has been exacerbated by technical difficulties for the robust characterization of dispersal events, especially those involving movement over long-distances (long-distance dispersal, LDD). LDD is a characteristically extreme event of propagule movement in any plant population, typically occurring with an extremely low probability but potentially reaching an extremely long distance. Some progress has recently been made through the fast-paced implementation of molecular tools in ecological research labs and the availability of cutting-edge technology for biotelemetry applications. But much of the population geneticist and ecologist communities remains unaware of the state of the art in each other and likely under-appreciates their potential to validate and enrich dispersal studies \citep{Jones:2008il}. In particular, LDD events remain difficult to assess, both technically- with serious methodological problems for its reliable estimation- and conceptually. Our aim here is to review the LDD concept with a specific emphasis on dispersal of plant propagules (seeds and pollen), providing an extended definition that might be helpful in the robust quantification of LDD events.

@@

No explicit definition of what constitutes an LDD event exists. Previous approaches \citep[e.g., ][]{Nathan:2006aa,Schurr2009long} include both absolute and proportional definitions to characterize LDD events. This means providing information about the absolute distances moved by a given percentile of the events and/or providing data on the proportion of events exceeding a given distance threshold \citep{Nathan:2008is}. The exact proportional or absolute thresholds selected remain arbitrary, as no reference spatial frame is provided within the definition of LDD. This leaves the consideration of LDD as an extreme form of context-dependent phenomenon, strongly dependent upon the scale of the biological process studied \citep{Kinlan:2005fb} and of the specific organism considered. For example, \cite{Kinlan:2005fb} used a spatial reference frame to characterize LDD events of marine organisms, where sedentary adults and larvae differ enormously in the spatial scales of their dispersal \citep{DAloia:2013fc}. Therefore, any measure of extent and reach of LDD events requires reference to an explicit spatial frame or "local" scale \citep{Kinlan:2005fb}.

@@

Long-distance dispersal is currently one of the most debated topics in dispersal ecology; it defines the connectedness within the network of local populations and the possibilities for range expansion and successful colonization events. I propose a first demogenetically-based, operational definition of what a long-distance dispersal event actually is, and review existing empirical literature on distance thresholds from population and genetic perspectives. I also show how molecular tools have been used to identify the respective contributions of different animal species to the LDD portion of dispersal kernels of pollen and seeds by setting empirically-derived distance thresholds. Finally, I highlight potential applications of molecular markers beyond the quantification of just the dispersal distances that prevails in current studies, e.g., experimental approaches to assess dispersal limitation and Janzen-Connell effects.

@@

A second consideration in terms of spatial boundaries, with effects on dispersal patterns, is the genetic neighborhood area $N^b_e$, i.e., the spatial extent including a subset of panmictic individuals within a population \citep{Wright:1943aa,Wright:1946aa}. Thus, the $N^b_e$ area can be equal to the whole extent of the population whenever the population is unstructured and there is evidence for random mating events among all the individuals. However, most populations and stands of long-lived trees show highly aggregated and clumped distributions \citep{Seidler:2006hx}, where relatively long distances may separate groups of individuals within the same population. In these cases we might expect $N^b_e$ area to be substantially smaller than the total population area. Therefore, at least four possible scenarios exist with distinct implications in terms of consequences for dispersal (Table 1). In the case of dispersal events not extending beyond the geographic limits of the population or reference area, actual LDD events may involve dissemination beyond a reduced neighborhood area that is smaller than the geographic extent of the population, originating local long-distance ($LDD_{loc}$) dispersal events (Table 1, Fig. 2A). Actual short-distance dispersal would then involve those situations where the propagule is disseminated within \textit{both} the population limits and the genetic neighborhood boundary ($SDD_{loc}$). Along a similar reasoning, dispersal events outside the population limits will not necessarily convey LDD (Table 1, Fig. 2B): this is expected in cases where the genetic neighborhoods are extensive, going beyond the geographic limits of local populations, as in fig trees \citep{Nason:1998aa}, generating LDD events within the genetic neighborhood ($LDD_{neigh}$). Finally, strict-sense LDD events would involve dissemination outside \textit{both} the population limits and the genetic neighborhood boundary ($LDD_{ss}$) (Table 1, Fig. 2A).

>>>>

The movement of pollen and seeds by animals and its consequences have intrigued population geneticists and field ecologists since the infancy of both research disciplines. Each has generated an impressive body of theoretical and empirical research through the past decades, yet advances have long been co-existing in ‘parallel worlds’ and the great synergistic potential of population genetics and demography for the study of plant dispersal by animals remains little explored. Knowledge gaps still having the imprint of this conceptual disconnection include the idea of long distance dispersal, and the paradoxes of forest fragmentation effects on genetic diversity \citep{Kramer:2008kg}, survival and persistence of relict tree species \citep{Hampe:2011bv}, rapid post-Pleistocene recolonization of vast continental areas in response to climate modification \citep{Clark:1998aa,Clark:1998vi}, among other persisting issues. This conceptual isolation has been exacerbated by technical difficulties for the robust characterization of dispersal events, especially those involving movement over long-distances (long-distance dispersal, LDD). Some progress has recently been made through the fast-paced implementation of molecular tools in ecological research labs and the availability of cutting-edge technology for biotelemetry applications. But much of the population geneticist and ecologist communities remains unaware of the state of the art in each other and likely under-appreciates their potential to validate and enrich dispersal studies \citep{Jones:2008il}. In particular, LDD events remain difficult to assess, both technically- with serious methodological problems for its reliable estimation- and conceptually. Our aim here is to review the LDD concept with a specific emphasis on dispersal of plant propagules (seeds and pollen), providing an extended definition that might be helpful in the robust quantification of LDD events.

@@

No explicit definition of what constitutes an LDD event exists. LDD is a characteristically extreme event of propagule movement in any plant or animal population, typically occurring with an extremely low probability but potentially reaching an extremely long distance. Previous approaches \citep[e.g., ][]{Nathan:2006aa,Schurr2009long} include both absolute and proportional definitions to characterize LDD events. This means providing information about the absolute distances moved by a given percentile of the events and/or providing data on the proportion of events exceeding a given distance threshold, i.e., an operational definition \citep{Nathan:2008is}. The exact proportional or absolute thresholds selected remain somehow arbitrary, as no reference spatial frame is provided within the definition of LDD. This leaves the consideration of LDD as an extreme form of context-dependent phenomenon, strongly dependent upon the scale of the biological process studied \citep{Kinlan:2005fb} and of the specific organism considered. For example, \cite{Kinlan:2005fb} used a spatial reference frame to characterize LDD events of marine organisms, where sedentary adults and larvae differ enormously in the spatial scales of their dispersal \citep{DAloia:2013fc}. Therefore, any measure of extent and reach of LDD events requires reference to an explicit spatial frame or "local" scale \citep{Kinlan:2005fb,Byrne:2014gu}.

@@

Long-distance dispersal is currently one of the most debated topics in dispersal ecology; it defines the connectedness within the network of local populations and the possibilities for range expansion and successful colonization events. I propose a first demogenetically-based, operational definition of what a LDD event actually is, and review existing empirical literature on distance thresholds from population and genetic perspectives. I also show how molecular tools have been used to identify the respective contributions of different animal species to the LDD portion of dispersal kernels of pollen and seeds by setting empirically-derived distance thresholds. Finally, I highlight potential applications of molecular markers beyond the quantification of just the dispersal distances that prevails in current studies, e.g., experimental approaches to assess dispersal limitation and Janzen-Connell effects.

@@

A second consideration in terms of spatial boundaries, with effects on dispersal patterns, is the genetic neighborhood area $N^b_e$, i.e., the spatial extent including a subset of panmictic individuals within a population \citep{Wright:1943aa,Wright:1946aa}. Thus, the $N^b_e$ area can be equal to the whole extent of the population whenever the population is unstructured and there is evidence for random mating events among all the individuals. However, most populations and stands of long-lived trees show highly aggregated and clumped distributions \citep{Seidler:2006hx}, where relatively long distances may separate groups of individuals within the same population. In these cases we might expect $N^b_e$ area to be substantially smaller than the total population area. Therefore, at least four possible scenarios exist with distinct implications in terms of consequences for dispersal (Table 1). In the case of dispersal events not extending beyond the geographic limits of the population or reference area, actual LDD events may involve dissemination beyond a reduced neighborhood area that is smaller than the geographic extent of the population, originating local long-distance ($LDD_{loc}$) dispersal events (Table 1, Fig. 2A). Actual short-distance dispersal would then involve those situations where the propagule is disseminated within \textit{both} the population limits and the genetic neighborhood boundary ($SDD_{loc}$). Along a similar reasoning, dispersal events outside the population limits will not necessarily convey LDD (Table 1, Fig. 2B): this is expected in cases where the genetic neighborhoods are extensive, going beyond the geographic limits of local populations, as in fig trees \citep{Nason:1998aa} with long-distance pollination, generating LDD events within the genetic neighborhood ($LDD_{neigh}$). Note that pollen and seeds may have contrasting movement patterns in rference to the distinct spatial scales of the population limits and of the genetic neighbourhood. For example, wind-dispersed species with reduced seed mobility (in terms of distance), such as oaks, can have large genetic neighbourhoods with extensive pollen dispersal \citep{Streiff99} \citep[but see, e.g.,][for fragmented stands]{Smouse:2001aa,Dutech:2005fk} so that $LDD_{neigh}$ dispersal events might frequently move beyond the physical limits of the population, patch, or stand but remain within the genetic neighbourhood. Finally, strict-sense LDD events would involve dissemination outside \textit{both} the population limits and the genetic neighborhood boundary ($LDD_{ss}$) (Table 1, Fig. 2A).

### 2016-10-10 Manuscript/SupplMat.tex
Minor edits.
Minor edits to title page.
Edited refs and Suppl Mat.

\title{Manuscript Draft\\
\vspace{2cm}
What is long-distance dispersal? And a taxonomy of dispersal events \\
\textbf{Supplementary Material}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Supplementary Material
%------------------------------------------------------------------- Table S1
\begin{landscape}
\begin{table}
  \caption*{Table S1. Summary of neighborhood area sizes and estimated neighborhood radius for tree species with different combinations of dispersal modes. Data from \citet{Nason:1998aa,Garcia:2005fu,Garcia:2007he} and present study.}
    \begin{tabular}{lllccc}
     Species                & Pollinator  & Seed disperser & Density ($ha^{-1}$) & Breeding unit ($km^2$) & Radius (km) \\
    \textit{Ficus dugandii}          & Fig wasp    & Vertebrates    & 0.004          & 631.7               & 14.2        \\
    \textit{Ficus obtusifolia}       & Fig wasp    & Vertebrates    & 0.072          & 105.9               & 5.8         \\
    \textit{Prunus mahaleb}          & Bees, flies & Vertebrates    & 0.003          & 0.87                & 0.042       \\
    \textit{Frangula alnus}          & Bees, flies & Vertebrates    & 0.0004         & 0.45                & 0.013       \\
    \textit{Astrocaryum mexicanum}   & Beetle      & Vertebrates    & 1364.0         & 0.011               & 0.06        \\
    \textit{Calophyllum longifolium} & Bees        & Vertebrates    & 0.28           & 1.241               & 0.629       \\
    \textit{Platypodium elegans}     & Bees        & Wind           & 0.78           & 0.866               & 0.525       \\
    \textit{Cedrus atlantica}        & Wind        & Wind           & 61.7           & 0.151               & 0.22        \\
   \textit{Fraxinus americana}      & Wind        & Wind           & 24.7           & 0.008               & 0.05        \\
    \textit{Pseudotsuga menziesii}   & Wind        & Wind           & 25.0           & 0.078               & 0.158       \\
    \end{tabular}
\end{table}
\end{landscape}
\newpage

>>>>

\title{\textbf{Supplementary Material}\\
\vspace{2cm}
What is long-distance dispersal? And a taxonomy of dispersal events \\
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Supplementary Material
%------------------------------------------------------------------- Table S1
\begin{landscape}
\begin{table}
  \caption*{Table S1. Summary of neighborhood area sizes and estimated neighborhood radius for tree species with different combinations of dispersal modes. Data from \citet{Nason:1998aa,Smouse:2001aa,Garcia:2005fu,Garcia:2007he,Dutech:2005fk} and present study.}
  \vspace{0.5cm}
    \begin{tabular}{lllccc}
        \hline \\
     Species                & Pollinator  & Seed disperser & Density ($ha^{-1}$) & Breeding unit ($km^2$) & Radius (km) \\\\ \hline \\
    \textit{Ficus dugandii}          & Fig wasp    & Vertebrates    & 0.004          & 631.7               & 14.2        \\
    \textit{Ficus obtusifolia}       & Fig wasp    & Vertebrates    & 0.072          & 105.9               & 5.8         \\
    \textit{Prunus mahaleb}          & Bees, flies & Vertebrates    & 0.003          & 0.87                & 0.042       \\
    \textit{Frangula alnus}          & Bees, flies & Vertebrates    & 0.0004         & 0.45                & 0.013       \\
    \textit{Astrocaryum mexicanum}   & Beetle      & Vertebrates    & 1364.0         & 0.011               & 0.06        \\
    \textit{Calophyllum longifolium} & Bees        & Vertebrates    & 0.28           & 1.241               & 0.629       \\
    \textit{Platypodium elegans}     & Bees        & Wind           & 0.78           & 0.866               & 0.525       \\
    \textit{Quercus lobata}          & Wind        & Vertebrates    & 3.0
      & 0.072               & 0.719       \\ 
    \textit{Quercus alba}            & Wind        & Vertebrates    & 92.8
      & 0.0009              & 0.017       \\ 
    \textit{Cedrus atlantica}        & Wind        & Wind           & 61.7           & 0.151               & 0.22        \\
   \textit{Fraxinus americana}       & Wind        & Wind           & 24.7           & 0.008                & 0.05        \\
    \textit{Pseudotsuga menziesii}   & Wind        & Wind           & 25.0           & 0.078               & 0.158       \\\\
\hline
    \end{tabular}
\end{table}
\end{landscape}
\newpage
